Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-03-04 src/tex/ms.tex
make bold, add additional variables
motivate energy flux instead of photon number flux

The goal of this work is to evaluate the potential of future exoplanet surveys to test the hypothesis that a minimum past \gls{NUV} flux is required for abiogenesis.
We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured to be $F_\mathrm{NUV, min} = \SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated from \SIrange{200}{280}{\nano\meter} at the surface~\citep{Rimmer2018,Rimmer2021a,Rimmer2023,Ranjan2023a}.
%\todo[inline]{SR: This is an interesting number, because it is below what was available on early Earth (so this scenario could have worked on early Earth) but until recently it was below what was thought to be available on \gls{HZ} M-dwarf exoplanets. So it was thought that identification of biosignatures on M-dwarf planets could therefore falsify the cyanosulfidic scenario, with a potential caveat for transient UV from flares. Two recent developments have complicated the picture. First, Rimmer et al. 2018 had an error in their radiative transfer routines. Correcting for this error, early M-dwarfs and highly active M-dwarfs emit enough UV to meet the Rimmer et al. 2018 criterion (Ranjan et al. 2023).
%\sout{Second, a recent publication argues that /all/ estimates of M-dwarf UV are underestimates, and that late M-dwarf stars have similar emission to Sunlike stars (Rekhi et al. 2023). I suspect this is incorrect, because it contradicts a lot of work from e.g. the MUSCLES collaboration and the HAZMAT project, but it's worth keeping an eye on in case it is correct after all.}}

We follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the \rev{\gls{NUV}} irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$, if \gls{NUV} exceeds the threshold for a sufficiently long period of time \rev{$\Delta T_\mathrm{min}$.}
Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the \gls{UV} flux.
\rev{
While we adopt $\Delta T_\mathrm{min} = \SI{1}{\mega\year}$, we find that increasing this timescale to \SI{100}{\mega\year} reduces the fraction of inhabited planets under $H_1$ by less than 10 \%, suggesting that our results are not highly sensitive to $\Delta T_\mathrm{min}$.
}

@@

\appendix
\section{Alternative priors for the abiogenesis rate}
\label{sec:appendix-flife-prior}
\rev{In the main text, we used a log-uniform prior for the probability of life emerging and persisting $f_\mathrm{life}$ in order to reflect our ignorance about this parameter and its order of magnitude.
We consider this a reasonable choice, as it is agnostic about the scale of $f_\mathrm{life}$ and assigns equal prior probability to all orders of magnitude.
Nevertheless, we repeated our semi-analytical analysis for the more optimistic choice of a prior that is \textit{uniform} in $f_\mathrm{life}$.
Figure~\ref{fig:semian_true_evidence_uniform} shows the resulting probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis under this prior.
While the resulting trends are qualitatively similar, the uniform prior yields overall higher probabilities of such a conclusive test for all sample sizes.
}

>>>>

The goal of this work is to evaluate the potential of future exoplanet surveys to test the hypothesis that a minimum past \gls{NUV} flux is required for abiogenesis.
We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured \rev{to be $\SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated} from \SIrange{200}{280}{\nano\meter} at the surface~\citep{Rimmer2018,Rimmer2021a,Rimmer2023,Ranjan2023a}.
%\todo[inline]{SR: This is an interesting number, because it is below what was available on early Earth (so this scenario could have worked on early Earth) but until recently it was below what was thought to be available on \gls{HZ} M-dwarf exoplanets. So it was thought that identification of biosignatures on M-dwarf planets could therefore falsify the cyanosulfidic scenario, with a potential caveat for transient UV from flares. Two recent developments have complicated the picture. First, Rimmer et al. 2018 had an error in their radiative transfer routines. Correcting for this error, early M-dwarfs and highly active M-dwarfs emit enough UV to meet the Rimmer et al. 2018 criterion (Ranjan et al. 2023).
%\sout{Second, a recent publication argues that /all/ estimates of M-dwarf UV are underestimates, and that late M-dwarf stars have similar emission to Sunlike stars (Rekhi et al. 2023). I suspect this is incorrect, because it contradicts a lot of work from e.g. the MUSCLES collaboration and the HAZMAT project, but it's worth keeping an eye on in case it is correct after all.}}
\rev{While prebiotic photochemistry is fundamentally driven by photon number flux, we choose to frame the problem in terms of energy flux, for which empirical estimates in the \gls{NUV} exist across stellar populations and ages~\citep{Richey-Yowell2023}.
In these units, the above threshold is approximately \var{rimmer_threshold}$\pm$\var{rimmer_uncertainty}$\,$\SI{}{\erg\per\second\per\centi\meter\squared}.
However, due to uncertainties in how surface flux relates to top-of-atmosphere flux in the habitable zone, we treat the threshold \gls{NUV} flux as a free parameter.}

We \rev{first} follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against \rev{the UV Threshold Hypothesis}.
Under \rev{this hypothesis} ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the \rev{\gls{NUV}} irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$, if \gls{NUV} exceeds the threshold for a sufficiently long period of time \rev{$\Delta T_\mathrm{min}$.}
Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the \gls{UV} flux.
\rev{
While we adopt $\Delta T_\mathrm{min} = \SI{1}{\mega\year}$, we find that increasing this timescale to \SI{100}{\mega\year} reduces the fraction of inhabited planets under $H_1$ by less than 10 \%, suggesting that our results are not highly sensitive to $\Delta T_\mathrm{min}$.
}

@@

\appendix
\section{\rev{Alternative priors for the abiogenesis rate}}
\label{sec:appendix-flife-prior}
\rev{In the main text, we used a log-uniform prior for the probability of life emerging and persisting $f_\mathrm{life}$ in order to reflect our ignorance about this parameter and its order of magnitude.
We consider this a reasonable choice, as it is agnostic about the scale of $f_\mathrm{life}$ and assigns equal prior probability to all orders of magnitude.
Nevertheless, we repeated our semi-analytical analysis for the more optimistic choice of a prior that is \textit{uniform} in $f_\mathrm{life}$.
Figure~\ref{fig:semian_true_evidence_uniform} shows the resulting probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis under this prior.
While the resulting trends are qualitatively similar, the uniform prior yields overall higher probabilities of such a conclusive test for all sample sizes.
}
